---
title: " SDS 315 Homework 9"
date: "2024-01-21"
output:
  pdf_document:
    toc: yes
---

Name: Neena Varanasi

EID: nv6558

[GitHub Link](https://github.com/Neena05/SDS_315_HW9)

```{r, message = FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(knitr)
library(mosaic)
library(MatchIt)
library(moderndive)
```

## Problem 1: Get out the vote
```{r, echo=FALSE, message = FALSE}
#read the dataset
turnout = read_csv("turnout.csv")

```

### Part A
```{r, echo=FALSE, message = FALSE}
#Part A
#The proportion of those receiving a GOTV call who voted in 1998.
prop_received = prop(voted1998 ~ GOTV_call, data=turnout, success=1)
#prop_received[2]

#The sample proportion of those not receiving a GOTV call who voted in 1998
#A large-sample 95% confidence interval for the difference in these two proportions
boot_1998 = do(1000)*diffprop(voted1998==1 ~ GOTV_call, data=resample(turnout))
ci = confint(boot_1998, level=0.95)

kable(confint(boot_1998, level=0.95))

```
The proportion of those receiving a GOTV call who voted in 1998 was `r round(prop_received[2], 2)`.

The sample proportion of those not receiving a GOTV call who voted in 1998 was `r round(prop_received[1], 2)`.

The 95% confidence interval for the difference in the two proportions is `r round(ci[2], 2)` to `r round(ci[3], 2)`.

### Part B
```{r, echo=FALSE, message = FALSE}
#Part B
summary_table <- turnout %>%
  mutate(GOTV_call_label = ifelse(GOTV_call == 1, "Recieved", "Did Not Recieve")) %>% 
  group_by(GOTV_call_label) %>%
  summarize(mean_age = mean(AGE), mean_majorpty = mean(MAJORPTY), mean_voted_1996 = mean(voted1996)) %>% 
  kbl(col.names = c("Get Out the Vote Call","Mean Age", "Mean Party Registration", "Voted in 1996"), booktabs = TRUE, digits = 2) %>%
  kable_styling(latex_options = "hold_position")
  
summary_table

ggplot(turnout) + 
  geom_boxplot(aes(x=factor(GOTV_call), y=AGE), fill="lightblue") + 
  coord_flip() +
  ggtitle("Age and Get Out the Vote Call") +
  ylab("Age (Years)") +
  xlab("Reciepient of Get Out the Vote Call") +
  theme_classic()
```

### Part C
```{r, echo=FALSE, message = FALSE}
#matching
turnout_matched = matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, data = turnout, ratio=5) %>%
  match.data

#prove confounders no longer exist
summary_table_matched <- turnout_matched %>%
  mutate(GOTV_call_label = ifelse(GOTV_call == 1, "Recieved", "Did Not Recieve")) %>% 
  group_by(GOTV_call_label) %>%
  summarize(mean_age = mean(AGE), mean_majorpty = mean(MAJORPTY), mean_voted_1996 = mean(voted1996)) %>% 
  kbl(col.names = c("Get Out the Vote Call","Mean Age", "Mean Party Registration", "Voted in 1996"), booktabs = TRUE, digits = 2) %>%
  kable_styling(latex_options = "hold_position")

summary_table_matched

#Repeat Part A statistics
prop_voted_2 = prop(voted1998 ~ GOTV_call, data=turnout_matched, success=1)


boot_1998_matched = do(10000)*diffprop(voted1998==1 ~ GOTV_call, data=mosaic::resample(turnout_matched))
kable(confint(boot_1998_matched, level=0.95))

prop.test(voted1998 ~ GOTV_call, data=turnout_matched)
```
The proportion of those receiving a GOTV call who voted in 1998 was `r round(prop_voted_2[2], 2)`.

The sample proportion of those not receiving a GOTV call who voted in 1998 was `r round(prop_voted_2[1], 2)`.

What do you conclude about the overall effect of the GOTV call on the likelihood of voting in the 1998 election?
Due to the 95% confidence interval being between `r round(confint(boot_1998_matched, level=0.95)[2] * 100, 2)`% and `r round(confint(boot_1998_matched, level=0.95)[3] * 100, 2)`%, we can conclude that GOTV likely does not have an effect on the likelihood of voting in the 1998 election. This conclusion is further supported by the p-value of `r round(prop.test(voted1998 ~ GOTV_call, data=turnout_matched)$p.value, 2)`, which is statistically significant since it is less than 0.05.


## Problem 2: Manufacturing flaws in circuit boards
```{r, echo=FALSE, message = FALSE}
#import the dataset
solder = read_csv("solder.csv")
```

### Part A
```{r, echo=FALSE, message = FALSE}
ggplot(solder) + 
  geom_boxplot(aes(x=(skips), y=Opening, fill = Opening)) + 
  coord_flip() +
  ggtitle("Opening Size of Solder and Number of Skips") +
  ylab("Opening Size of Solder (small, medium, or large)") +
  xlab("Number of Solder Skips") +
  theme_classic()
```
Based on the plot, the smaller opening sized solders have a higher number of skips compared to medium and large sized opening solders. Medium solders still have a higher number of skips than large solders.

```{r, echo=FALSE, message = FALSE}
ggplot(solder) + 
  geom_boxplot(aes(x=(skips), y=Solder, fill=Solder)) + 
  coord_flip() +
  ggtitle("Age and Get Out the Vote Call") +
  ylab("Thickness of the Alloy Used (thick or thin)") +
  xlab("Number of Solder Skips") +
  theme_classic()
```
Based on the plot, the thin alloy solders have a higher number of skips compared to thick alloy solders.

```{r, echo=FALSE, message = FALSE}
solder %>%
  group_by(Opening) %>%
  summarize(skips_proportion = sum(skips) / sum(solder$skips)) %>% 
  kbl(col.names = c("Opening Size","Percent of Solder Guns with Manufacturing Flaws"), booktabs = TRUE, digits = 2) %>%
  kable_styling()

solder %>%
  group_by(Solder) %>%
  summarize(skips_proportion = sum(skips) / sum(solder$skips)) %>% 
  kbl(col.names = c("Solder Thickness","Percent of Solder Guns with Manufacturing Flaws"), booktabs = TRUE, digits = 2) %>%
  kable_styling()

```
The above tables specifically show the quantitative difference between opening size and number of skips and solder thickness and number of skips. 


### Part B
```{r, echo=FALSE, message = FALSE}
#Part B
lm_opening = lm(skips ~ Opening, data=solder)
#summary(lm_opening)
#coef(lm_opening) %>% round(0)


lm_solder_type = lm(skips ~ Solder, data=solder)
#summary(lm_solder_type)
#coef(lm_solder_type) %>% round(0)

lm_opening_type = lm(skips ~ Opening + Solder + Opening:Solder, data=solder)
#summary(lm_opening_type)
kable(coef(lm_opening_type) %>% round(0))

#print regression table
kable(get_regression_table(lm_opening_type, conf.level = 0.95, digits=2))

```
### Part C

The baseline number of skips for solder guns that neither have a large opening and a thick alloy is 0 skips.
The main effect for the OpeningM variable is 2 skips. Therefore, having a medium opening is associated with an increase of 2 skips compared to a large opening in isolation.

The main effect for the OpeningS variable is 5 skips. Therefore, having a small opening is associated with an increase of 5 skips compared to a large opening in isolation.

The main effect for the SolderThin variable is 2 skips. Therefore, having a thin alloy is associated with an increase of 2 skips compared to a thick alloy in isolation.
The interaction effect for OpeningM and SolderThin is -1. In other words, having the effect of a medium opening and thin alloy decreases the number of skips by 1 than what you would expect from summing the individual "isolated" effects of the two variables.

The interaction effect for OpeningS and SolderThin is 10. In other words, having the effect of a small opening and thin alloy increases the number of skips by 10 than what you would expect from summing the individual "isolated" effects of the two variables..


### Part D
In order to minimize the number of skips, I would recommend a medium opening and a thin alloy thickness as the interaction effect demonstrated 1 fewer skips than what you would expect from a large, thick solder in isolation.
